---
title: "Laura_attempts_holy_grail_figure"
author: "Laura Bogar"
date: "3/24/2020"
output: html_document
---

I would like to build on the work that Kristin Klitgaard did to build the "Holy Grail" figure that Holly described:
P (plant biomass) will be on the y axis, with the proportion of time spent in env. A vs env. B on the x axis (all A, 50/50, all B). There will be three lines plotted, representing plants with just fungus A, plants with just fungus B, and plants with both fungi. Line A will start high in the all A environment, and end low; B will do the opposite. The tree with two fungi will show a hump in the middle (variable environment is best).

I would like to:
1) Make seasons longer so the curves are smoother. (New 400 value is not taking).
2) Optimize plant allocation in single-season, two-fungus situations to parameterize changing seasons.

Kristin did a great job setting up the following definitions:

State Variables 

P = individual plant biomass (tree)
A = individual fungus A biomass
B = individual fungus B biomass
N_P = Nutrient pool in the plant
N_S = Nutrient pool in the soil

Parameters

g = growth rate of tree
s = senescence. density independent mortality of tree (loss of leaves, branches, etc)
d_P = density dependent mortality of tree (as a tree gets bigger, more likely to get hit by lightning etc.)
r_A = rewards (in C) of tree to fungus. cost for tree, gain for fungus. 
e_A = efficiency of conversion for tree C to fungus C. 
m_A = density independent mortality of fungus
d_A = density independent mortality of fungus
u_P = nutrient uptake rate of tree 
u_A = nutrient uptake rate of fungus
l = leakiness of tree. nutrients that leave the tree and reenter the soil pool (decomposed leaves, root leakiness, etc.)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load required packages
require(tidyverse)
```
Parameter Set

NOTES TO LAURA:
- efficiency of 0.1 is so low! Is this right for fungi munching on tree sugars?

```{r}
g = 5 #growth rate of tree is higher than fungus
s = 1 #senescence is low compared to growth rate
e_A = .1 #standard energy flow between two trophic levels
e_B = .1 #standard energy flow between two trophic levels
r_A = 0.4 # 0.8 was close to optimal investment for single fungus (see IV.)
r_B = 0.4 # NOT BASED ON ANYTHING RIGHT NOW.
m_A = 0 #set to zero so ZNGI goes through origin (see III.)
m_B = 0
u_A = 1 #uptake of fungus is greater than plant to provide benefit ... ensure mutualism
u_B = 1
u_P = 0.01 #uptake of plant lower than fungus 
l = 1 #small compared to total pool
d_A = 1 #easy value
d_B = 1.5 # greater density-independent mortality than A
d_P = 0 #set to zero to prevent equations from being insane
N_T = 5 #initial soil nutrient pool level
```

Arbitrary initial values:
```{r}
timespan = 2000 # If we say each time point is one day, two thousand days will allow us to observe a bit more than 5 years of tree decision-making
P_init = 0.1 # initialize biomass as non-zero small number
A_init = 0.1
B_init = 0.1
Np_init = 0.1 # arbitrary initial nitrogen concentration for plant
```
Setting up the simulations:
```{r}
timepoints = 0:timespan #create a time vector going from 0 to the maximum time span we want to examine.
P.simu = NaN*timepoints; P.simu[1] = P_init # initialize simulated plant biomass
A.simu = NaN*timepoints; A.simu[1] = A_init # initialize simulated fungus A biomass
B.simu = NaN*timepoints; B.simu[1] = B_init
Np.simu = NaN*timepoints; Np.simu[1] = .1 #relatively arbitrary start number. 
Ns.simu = NaN*timepoints; Ns.simu[1] = N_T - Np.simu[1] #Nutrients in plant (Np) and nutrients in soil (Ns) add to total nutrients in system (N_T)
```

Simulate plant and fungal biomass in the absence of seasons:
```{r}
for(i in 2:length(timepoints)){ #for loop with the equations for N, P, Np, and Ns within the model. 
  dt = timepoints[i] - timepoints[i-1]
  P = P.simu[i-1]
  A = A.simu[i-1]
  B = B.simu[i-1]
  N_P = Np.simu[i-1]
  N_S = Ns.simu[i-1]
  dP = (P* (g*N_P/P - r_A*A - s*(1+d_P*P)))*dt
  dA = (A* (e_A*r_A*P-m_A-d_A*A))*dt
  dB = (B* (e_B*r_B*P - m_B-d_B*B))*dt
  dNp = (u_P*P*N_S + u_A*A*N_S - l*N_P)*dt
  dNs = (-u_P*P*N_S - u_A*A*N_S + l*N_P)*dt
  P.simu[i] = P + dP
  A.simu[i] = A + dA
  B.simu[i] = B + dB
  Np.simu[i] = N_P + dNp
  Ns.simu[i] = N_S + dNs
  
}

# And remodel the data to make plotting easier:
simulation_results = data.frame(time = timepoints, Plant = P.simu, Fungus_A = A.simu, Fungus_B = B.simu)

longform = simulation_results %>% gather(organism, "Biomass", -time)

# Plot it
ggplot(longform, aes(x = time, y = Biomass, group = organism, color = organism)) +
  geom_line()

```

Great! It seems to be working.

Let's add in seasons!
```{r Adding seasons}
seasonlength = 400 # let's say each season should last 100 days
badseasonfactor = 1.5 # mortality in bad season will be this times 
# the good season mortality. Currently same factor for both fungi.

season = rep(c(rep("A", seasonlength), rep("B", seasonlength)), ceiling((max(timepoints)/seasonlength)))

d_A.var = NaN*timepoints
d_B.var = NaN*timepoints
for (i in 1:length(timepoints)) {
  if (season[i] == "A") {
    d_A.var[i] = d_A
    d_B.var[i] = d_B*badseasonfactor
  } else if (season[i] == "B") {
    d_A.var[i] = d_A*badseasonfactor
    d_B.var[i] = d_B
  }
}

# d_A.variable = rep(c(rep(d_A, seasonlength), rep(d_A*badseasonfactor, seasonlength)), ceiling((max(timepoints)/seasonlength)))
# # This will generate slightly more d_A.e values than we need
# # if we end up with a timepoints value that's not evenly 
# # divisible by season, but this shouldn't be a problem for our 
# # purposes. We can just ignore them!
# 
# d_B.variable = rep(c(rep(d_B*badseasonfactor, seasonlength), rep(d_B, seasonlength)), ceiling((max(timepoints)/seasonlength)))

# Below, find Kristin's slightly more elegant solution:
# for(i in 1:length(tset)){ #for loop to vary the values of d_A by season
#   time <- tset[i]
#   if(floor(time/switchfreq) %% 2 == 0){ #divide the current timestep by the switching frequency, then round it. if this value when divided by 2 has no remainder, then you find yourself in season 1, and d_A should be smaller. 
#     d_A.e[i] <- d_A #d_A is the previously set value of 1
#   } else{ #otherwise, you are in season 2!
#     d_A.e[i] <- 1.5 #d_A is 50% greater, value of 1.5. value chosen to demonstrate difference. 
#   }
# }

```
```{r Simulation with two fungi and seasons}
# Here, I am simply varying fungal mortality with season.
# Kristin did something more sophisticated,
# shifting plant allocation to each fungus
# based on the season by finding optimal
# single-fungus values in each condition.

for(i in 2:length(timepoints)){
  dt = timepoints[i] - timepoints[i-1]
  P = P.simu[i-1]
  A = A.simu[i-1]
  B = B.simu[i-1]
  N_P = Np.simu[i-1]
  N_S = Ns.simu[i-1]
  dP = (P* (g*N_P/P - r_A*A - s*(1+d_P*P)))*dt
  dA = (A* (e_A*r_A*P-m_A-d_A.var[i]*A))*dt
  dB = (B* (e_B*r_B*P - m_B-d_B.var[i]*B))*dt
  dNp = (u_P*P*N_S + u_A*A*N_S - l*N_P)*dt
  dNs = (-u_P*P*N_S - u_A*A*N_S + l*N_P)*dt
  P.simu[i] = P + dP
  A.simu[i] = A + dA
  B.simu[i] = B + dB
  Np.simu[i] = N_P + dNp
  Ns.simu[i] = N_S + dNs
  
}

# And remodel the data to make plotting easier:
simulation_results = data.frame(time = timepoints, Plant = P.simu, Fungus_A = A.simu, Fungus_B = B.simu)

longform = simulation_results %>% gather(organism, "Biomass", -time)

# Plot it
ggplot(longform, aes(x = time, y = Biomass, group = organism, color = organism)) +
  geom_line()

```

This is a weird and spiky graph!